A meta-analysis of fungicide spray schedules to minimise yield loss from powdery mildew in Australian mungbean

### 2020-08-12 01_Meta-analysis_experiments.Rmd
Apply selection criteria to dataset and save variable to Rdata file

```{r MEsummary_Libraries, message=FALSE, echo=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, leaflet, htmltools, schoolmath, here)
source(here("R/import_data.R")
)
```
Trials testing fungicide efficacy on powdery mildew in mungbean in the Grains Research and Development Corporation (GRDC) northern grains region were sourced for a meta-analysis.

@@

   1. A field trial testing the efficacy of fungicide on powdery mildew afflicted mungbean plants.  
   2. Trial data needed to include:  
      i) fungicide active ingredients.  
      ii) A Demethylase inhibitor (DMI) fungicide.  
      iii) The data at which powdery mildew first appeared in the trial.  
      iv) Disease severity at the end of the growing seasons.  
      v) Fungicide application dates.  
      vi) fungicide dose.  
      vii) Crop yield.  
      viii) Treatment means and accompanying variance

@@

### Import data from trials we have the raw data  
```{r import_dataRAW}
PM_MB_raw <- 
   rbind(
      read.csv(here("cache/2010 PMmung Hermitage means.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Herm means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Kingaroy means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1305_Goolhi_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1304-MarysMount_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1303-Premer-Disease_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/BB1305_Millmerran_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Herm_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/King_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Fogerty_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Wellcamp_18_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_19_mean.csv"),stringsAsFactors = FALSE)
      )
write.csv(PM_MB_raw, file = here("cache/PM_Mungbean_SummaryOfTrialsWithRawData"), row.names = FALSE)
```
### Import dataset of manual entries  
Import dataset which were manually entered from trial reports
```{r import_dataMan}
source(here("R/import_ManData.R"))
PM_MB_man <- import_ManData()
```

@@

## Summary of all trials  
```{r data_summary, echo=FALSE, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   arrange(year, location) %>%
   group_by(trial_ref) %>%
   summarise(
         Year = unique(year),
         Location = unique(location),
         `Replicates per treatment` = if(all(replicates[1] == replicates)){
            as.character(replicates[1])}else{
               as.character(paste(min(replicates),"-",max(replicates)))},
         `Planting date` = unique(planting_date),
         `First sign of disease` = unique(first_sign_disease),
         `Fungicide treatments` = length(n_treatment),
      ) %>%
      arrange(Year)%>%
   kable(caption = "Description of Experiments",
         align = "c",
         col.names = c("Unique Trial\nReference","Year", "Location", "Replicates\nper treatments",
                       "Planting date", "First sign\nof disease", "Fungicide\n treatments")
         )  %>%
   kable_styling(fixed_thead = TRUE, full_width = TRUE) %>%
   column_spec(c(3,5:6), width = "3cm") %>% 
   scroll_box(height = "500px")

@@

## Powdery mildew rating scales

**2001 - 2004 trials**

Trial data were obtained from a non-published manuscript.

### 1 - 5 powdery mildew rating scale

Ratings for powdery mildew were on a 1 - 5 rating scale where

   1 - No sign of disease

   2 - 1 - 30&nbsp;% diseased leaf area

   3 - 31 - 65&nbsp;% diseased leaf area

   4 - 66 - 99&nbsp;% diseased leaf area and

   5 - Complete defoliation

Comparing these trials which have not used the 1 - 9 rating scale will be converted using a percentage change.

**2013 Field trials**

Trials from 2013 were all conducted by the [Northern Growers Alliance (NGA)](https://www.nga.org.au/).
The raw data were obtained from Lawrie Price.
NGA data were formatted to be compliant with the 1 - 9 rating scale used in all other experiments.
Detailed code for the conversion can be found in the following vignette [191115_PMMB_DataWrangling_PM.Rmd](./vignette/191115_PMMB_DataWrangling_PM.html).

### 1 - 9 powdery mildew rating scale

   1 - No powdery mildew colonies observed on any plant

   2 - Small colonies in lower 1/3 of canopy, up to 75&nbsp;% of plants affected

   3 - Colonies in lower 1/2 of canopy, >75&nbsp;% of plants affected

   4 - Colonies in lower 2/3 of canopy, up to 7&nbsp;5% of plants affected

   5 - Colonies in lower 2/3 of canopy, >75&nbsp;% of plants affected

   6 - Colonies in lower 2/3 of canopy, 100&nbsp;% of plants affected

   7 - Colonies in lower 2/3 of canopy of 10&nbsp;0% of plants, some plants with colonies also in top 1/3 of canopy

   8 - Colonies to top of plant with >75&nbsp;% of plants affected

   9 - Colonies to top of plant with 100&nbsp;% of plants affected and heavy leaf drop

>>>>

```{r MEsummary_Libraries, message=FALSE, echo=FALSE}
if (!require("pacman"))
   install.packages("pacman")
pacman::p_load(tidyverse, kableExtra, leaflet, htmltools, lubridate, here)

if (!require("theme.usq"))
   devtools::install_github("adamhsparks/theme.usq")
library(theme.usq)
theme_set(theme_usq())
```
Trials testing fungicide efficacy on powdery mildew in mungbean in the Grains Research and Development Corporation (GRDC) northern grains region were sourced for a meta-analysis.

@@

   1. A field trial testing the efficacy of fungicide on powdery mildew afflicted mungbean plants.  
   2. Trial data needed to include:  
      i) fungicide active ingredients.  
      ii) A Demethylase inhibitor (DMI) fungicide.  
      iii) The date at which powdery mildew first appeared in the trial.  
      iv) Disease severity at the end of the growing seasons.  
      v) Fungicide application dates.  
      vi) fungicide dose.  
      vii) Crop yield.  
      viii) Treatment means and accompanying variance

@@

### Import data from trials we have the raw data  
```{r import_dataRAW}
PM_MB_raw <- 
   rbind(
      read.csv(here("cache/2010 PMmung Hermitage means.csv"), stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Herm means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/2011 PMmung Kingaroy means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1305_Goolhi_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1304-MarysMount_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/AM1303-Premer-Disease_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/BB1305_Millmerran_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Herm_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/King_16_means.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Fogerty_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_17_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Wellcamp_18_mean.csv"),stringsAsFactors = FALSE),
      read.csv(here("cache/Hermitage_19_mean.csv"),stringsAsFactors = FALSE)
      )
write.csv(PM_MB_raw, file = here("cache/PM_Mungbean_SummaryOfTrialsWithRawData"), row.names = FALSE)
```

### Import dataset of manual entries  
Import dataset which were manually entered from trial reports  
```{r import_dataMan}
PM_MB_man <- read.csv(here("cache/PM_Mungbean_DataManuallyEntered.csv"), stringsAsFactors = FALSE)
```

@@

### Summary of all trials  
```{r data_summary, echo=FALSE, message=FALSE, warning=FALSE}
PM_MB_dat %>%
   arrange(year, location) %>%
   group_by(trial_ref) %>%
   summarise(
         Year = unique(year),
         Location = unique(location),
         `Replicates per treatment` = if(all(replicates[1] == replicates)){
            as.character(replicates[1])}else{
               as.character(paste(min(replicates),"-",max(replicates)))},
         `Planting date` = unique(planting_date),
         `First sign of disease` = unique(first_sign_disease),
         `Fungicide treatments` = length(n_treatment),
      ) %>%
      arrange(Year)%>%
   kable(caption = "Description of Experiments",
         align = "c",
         col.names = c("Unique Trial\nReference","Year", "Location", "Replicates\nper treatments",
                       "Planting date", "First sign\nof disease", "Fungicide\n treatments")
         )  %>%
   kable_styling(fixed_thead = TRUE, full_width = TRUE) %>%
   column_spec(c(3,5:6), width = "3cm") %>% 
   scroll_box(height = "500px")

@@

Lets apply the selection criteria our dataset.  
All trials identified for this meta-analysis reported: fungicide active ingredient, dose, first sign of disease. 
Therefore no trials need to be removed to satisfy this criteria.

Three trials omitted due to not reporting yields  
```{r TrialWithoutYield}
PM_MB_dat %>%
   filter(is.na(grain_yield.t.ha.)) %>%
   distinct(trial_ref, year, location)

PM_MB_dat <- PM_MB_dat %>%
   filter(!is.na(grain_yield.t.ha.))
```
One more trial removed as it did not report disease severity.  
```{r TrialsWithotDisease}
PM_MB_dat %>%
   filter(is.na(PM_final_severity)) %>%
   distinct(trial_ref, year, location)

PM_MB_dat <- PM_MB_dat %>%
   filter(!is.na(PM_final_severity))
```

No trials from our subset need to be removed for not reporting fungicide application dates.  
```{r FungicideApplicationDates}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Application_Dates = all(is.na(fungicide_application_1)),
             .groups = 'drop')%>%
   filter(No_Record_Of_Fungicide_Application_Dates)
```
No trials from our subset need to be removed for not reporting fungicide dose.  
```{r FungicideDoseTrials}
PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(No_Record_Of_Fungicide_Dose = all(is.na(dose_ai.ha)),
             .groups = 'drop')%>%
   filter(No_Record_Of_Fungicide_Dose)
```

Four trials from our subset need to be removed for not reporting yield variance.  
```{r TrialsWithVariance}
NoVar <- PM_MB_dat %>%
   group_by(trial_ref) %>%
   summarise(location = location,
             year = year,
             No_Var = all(is.na(yield_error)),
             .groups = 'drop')%>%
   filter(No_Var)%>%
   distinct(trial_ref, year, location)
NoVar

PM_MB_dat <- PM_MB_dat %>%
   filter(!(trial_ref %in% NoVar$trial_ref))
```

We need to focus the meta-analysis on fungicides with the same mode of action.
Let's look at all the fungicides and retain the best represented fungicide group in the data.
```{r Fungicides}
PM_MB_dat %>%
   group_by(fungicide_ai, trial_ref) %>%
   summarise(groups = 'drop') %>%
   count(sort = TRUE) %>%
   rename(Trials = n) %>%
   ggplot(aes(x = reorder(fungicide_ai, Trials), y = Trials)) +
   xlab("Fungicide active ingredient") +
   ylab("N Trials") +
   geom_col() +
   scale_fill_usq() +
   ggtitle(label = "Number of trials in which the\nspecified fungicide was used") +
   scale_colour_usq() +
   coord_flip()
```

Tebuconazole and propiconazole are both classed as demethylase inhibitors. 
We will retain the trials trials which only contain DMI fungicides.  
```{r}
PM_MB_dat %>%
   filter(fungicide_ai == "tebuconazole" |
             fungicide_ai == "propiconazole")%>%
   distinct(trial_ref, location, year)%>%
   arrange(year)
```
We will retain all fungicides for now and remove them before the analysis.

Finally we are going to define the class of each variable in the data, retaining only the variables relevant to the analysis.
```{r classifyData}
PM_MB_dat <- 
   PM_MB_dat %>%
   mutate(trial_ref = as.character(trial_ref)) %>%
   mutate(year = as.character(year)) %>%
   mutate(location = as.character(location)) %>%
   mutate(host_genotype = as.character(host_genotype)) %>%
   mutate(row_spacing = as.double(round(row_spacing, digits = 2))) %>%
   mutate(replicates = as.integer(replicates)) %>%
   mutate(planting_date = ymd(planting_date)) %>%
   mutate(flowering_date = ymd(flowering_date)) %>%
   mutate(pod_fill_date = ymd(pod_fill_date)) %>%
   mutate(mid_late_pod_fill = ymd(mid_late_pod_fill)) %>%
   mutate(first_sign_disease = ymd(first_sign_disease)) %>%
   mutate(fungicide_ai = as.character(fungicide_ai))%>%
   mutate(dose_ai.ha = round(as.double(dose_ai.ha), digits = 2))%>%
   mutate(fungicide_application_1 = ymd(fungicide_application_1)) %>%
   mutate(fungicide_application_2 = ymd(fungicide_application_2)) %>%
   mutate(fungicide_application_3 = ymd(fungicide_application_3)) %>%
   mutate(fungicide_application_4 = ymd(fungicide_application_4)) %>%
   mutate(fungicide_application_5 = ymd(fungicide_application_5)) %>%
   mutate(fungicide_application_6 = ymd(fungicide_application_6)) %>%
   mutate(total_fungicide = as.integer(total_fungicide)) %>%
   mutate(harvest_date = ymd(harvest_date)) %>%
   mutate(final_assessment = ymd(final_assessment)) %>%
   mutate(PM_final_severity = round(as.double(PM_final_severity), digits = 4))%>%
   mutate(disease_error = round(as.double(disease_error), digits = 6))%>%
   mutate(D_error_type = as.character(D_error_type))%>%
   mutate(grain_yield.t.ha. = round(as.double(grain_yield.t.ha.), digits = 4))%>%
   mutate(yield_error = round(as.double(yield_error), digits = 6))%>%
   mutate(Y_error_type = as.character(Y_error_type))%>%
   mutate(Y_Msquare = round(as.double(Y_Msquare), digits = 6))%>%
   mutate(yield_gain = round(as.double(yield_gain), digits = 4))%>%
   mutate(prop_YG = round(as.double(prop_YG), digits = 4))%>%
   select(-c(trial_design, # remove the following columns
             plot_length.m.,
             plot_width.m.,
             plant_density,
             trade_name,
             n_treatment,
             emergence_date,
             rating_scale,
             raw_graded,
             raw_data_avail,
             comments,
             days_harvest.planting,
             AUDPS_m,
             AUDPC_sd,
             Inc_Ms))
```

```{r saveData01}
save(PM_MB_dat, file = here("cache/ImportDataAndSelectTrials01.Rdata"))
```
